---
title: "Makefile"
author: "Elizabeth Ramsdell"
date: "October 27, 2016"
output: 
  html_document:
  keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

This is a research project on the GDP and GDP ranking of different countries and how income group segments this. 

To do this, first, 2 data sets need to be merged in order to get all the variables necessary. Then, during exploratory analysis, I have looked at where certain countries rank among the others. I have also researched the average GDP rankings for the high income groups.

To visualize the results of GDP vs Income Group, a scatterplot and a table displaying the relationship has been included.


##Importing the datasets
```{r}
setwd("~/Doing Data Science/CaseStudy1/Data")
source("ImportData.R")

```

##Explore, Clean and Merge Data

```{r}
str(GDP)
str(Education)

```

US.dollars. is character and has commas. Change US.dollars. variable in GDP and rename 
```{r}
GDP$USDollars <- as.numeric(gsub(",","",GDP$US.dollars.))
```
Merge the GDP and Education Datasets and Save
Change all variable names to lower case
```{r}
MergeData <- merge(x=GDP, y=Education, by="CountryCode")

write.csv(MergeData,"MergeData")

names(MergeData) <- tolower(names(MergeData))

```

#1. How many of the IDs match? 
###189
```{r}
str(MergeData)
```

#2. Sort the data frame in ascending order by GDP, What is the 13th country in the resulting data frame?
###13th Country in ordered data is St. Kitts and Nevis
```{r}
OrderByGDP <- MergeData[order(MergeData$usdollars),]

OrderByGDP[1][13,]

KDPName <- subset(MergeData, countrycode == "KNA")
KDPName

```

#3. What are the average GDP rankings for the "High income: OECD" and "High income: nonOECD" groups?
###The average GDP rankings for the High Income OECD group is 32.96
###Average ranking for the High Income Non OECD group is 91.91
```{r}
unique(MergeData$income.group) ##Check for mispellings

HighIncomeOECD <- subset(MergeData, income.group == "High income: OECD")
mean(HighIncomeOECD$ranking)

HighIncomeNonOECD <- subset(MergeData, income.group == "High income: nonOECD")
mean(HighIncomeNonOECD$ranking)

```

## 5. Cut the GDP Ranking into 5 separate quantile groups
find ranges
```{r}
189/5
(189/5)*2
(189/5)*3
(189/5)*4
(189/5)*5
```
Create Categories
```{r}
attach(MergeData)
MergeData$rankingcat[ranking <= 37] <- "Low Rank"
MergeData$rankingcat[ranking > 37 & ranking <= 75] <- "Medium Low Rank"
MergeData$rankingcat[ranking > 75 & ranking <= 113] <- "Medium Rank"
MergeData$rankingcat[ranking > 113 & ranking <= 151] <- "Medium High Rank"
MergeData$rankingcat[ranking > 151] <- "High Rank"
detach(MergeData)



```

## Plots and Tables

#4. Plot GDP by Country

Table and Plot of GDP ranking and income group
```{r}
library(ggplot2)
ggplot(MergeData, aes(x=countrycode, y=log(usdollars))) + geom_point(aes(colour=income.group))

```

#5. Table:Ranking Category vs Income Group
How many countries are Lower middle income but among the 38 nations with highest GDP?
16
```{r}
table(MergeData$rankingcat,MergeData$income.group)
```


##Summary

####After merging and cleaning the 2 datasets, we are able to check any trends between GDP and Income Group. The scatter plot is colored by Income Group. Although the HighIncome:OECD group seems to have all of it's distribution near the top of the graph, the rest of the groups seem evenly spread throughout the graph. This might show that there is not an important relationship between income group and GDP rank within the countries given. The high income non OECD has a larger average than the high income OECD group. Also, the table outlines that the GDP rank may be spread throughout the different income groups, more than one would expect. 